\chapter{Experimental study}
\startcontents[chapters]

\section{Methodology}
We decided to test speakers' preferences for syntactic and semantic agreement in two online experiments: one using one type of lexical hybrids, profession nouns referring to women, the other using split hybrids, depreciative forms which have a shape of non-masculine nouns, but refer to men. In the experiments we tested predicative verb agreement and relative pronoun agreement.

I submitted the design of the experiments and the materials to the Paris Graduate School of Linguistics Ethics Committee and got their approval in May 2026. The following subsections describe the experiments in more detail.

\subsection{Procedure}
We conducted two acceptability judgement experiments: one using stimuli with profession nouns and the other with depreciative forms. Both experiments had the same design, which was 3x2 within-subject. The independent variables were the agreement options (semantic or syntactic) and type of agreement with ordering of target and controller: predicate with target first (prenominal), predicate with controller first (postnominal) and relative pronoun (always with controller first, since it is impossible to place the relative pronoun before its antecedent). The dependent variable was the rating of acceptability of given sentences.

The experiment was conducted in Polish. It was recommended to participants to complete the experiment using a computer or tablet, to make sure the instructions and stimuli were displayed correctly. On the first page participants were told their task was to judge whether the displayed sentences sounded naturally or not. They were also informed that participation in the study was voluntary and they could withdraw at any time. To proceed to the study they had to tick a box to confirm they consent. Next, they saw the instructions page, which had the following information.

{\slshape
  Your task will be to judge how naturally the displayed sentences sound. If you aren't sure what this means, consider whether you could hear or say the given sentence. You will judge the sentences on a 5-point scale, from ``completely unnatural'' to ``completely natural''. You will see a control question after each sentence.

  Before we proceed to the study, you will see 3 practice questions that will allow you to familiarize yourself with the task. After the practice session you will proceed to the study, which consists of 60 questions. During the study you cannot go back to the previous questions and there is no time limit to answer the questions, but answer intuitively, without thinking too much.
}

Out of 60 trials, 24 were with experimental items. They were arranged using the Latin square design -- in 6 lists, so that each item appears on a list once and each list has 4 items per condition. Each participant was assigned a list randomly and the order of items was randomized each time. Each trial was followed by a simple comprehension question, to ensure that participants are paying attention to the experiment.

In the experiment on profession nouns the instruction in each trial was ``Read the sentences below and judge how naturally the \textbf{\textsl{second}} one of them sounds'', for the experiment on depreciative forms it was ``Read the sentence below and judge how naturally it sounds''. As was said in the instructions for participants, judgements were given on a 5-point scale. Such a scale is familiar for Polish speakers, because it is similar to the grading scale in Polish schools (1-6 with 6 being the best) and in Polish universities (2-5 with 5 being the best). The scale was labelled only on the ends with \mean{całkowicie nienaturalne}{completely unnatural} and \mean{całowicie naturalne}{completely natural}.

After participants gave judgements on 60 sentences, there was a short demographic survey asking about their age, gender, first language(s) and the country or countries where they were raised. All of the response fields for those questions allowed for any kind of text answer to be typed. The duration of the whole experiment was 15-20 minutes.

\subsection{Materials}
Experimental items were constructed so that the agreement controller was either in the first or the final position in the sentence, the agreement target was immediately adjacent to the controller and there were no other words in the sentence that agreed with the controller in gender. To make sure that the last constraint was not violated, all of the sentences with relative pronoun agreement had to be in present tense (because predicates in present tense do not agree with the subject in gender), while the sentences in other conditions were in past tense. The following subsections describe how the nouns for the experimental items were selected.

\subsubsection{Experimental items with depreciative forms}
Depreciative forms appear in the paradigms of \textsc{m1} nouns and morphologically they resemble nouns of genders other than \textsc{m1}. In Corbett's approach then, when using those nouns in a sentence we have a syntactic reason to use non-\textsc{m1} agreement and a semantic reason to use \textsc{m1} agreement. To make sure that semantic reason is clearly pointing to \textsc{m1}, the nouns selected for this experiment had to be commonly perceived as masculine. To ensure that, the selection process relied on a norming study conducted by \cite{misersky_norms_2014}, where they collected information about gender stereotypicality for role nouns in multiple Eurpean languages.

There were 22 nouns that fulfilled the following criteria:
\begin{itemize}
  \item Perceived to be at least 60\% masculine according to the norming study.
  \item Appear at least a 1000 times in the balanced version of the National Corpus of Polish.
  \item Depreciative form is not syncretic with the non-depreciative.
  \item Does not have disturbing, violent connotations (e.g. rapist, murderer)
  \item Depreciative form is not syncretic with a different noun (e.g. noun \mean{technik}{technician} is \textsl{techniki} in depreciative, which has the same shape as the word for `techniques' and could lead to confusion)
  \item Is not a nominalized adjective (for which depreciative forms are syncretic with the feminine versions of these nouns)
\end{itemize}
Additionally, there were two nouns that did not occur in the norming study: \mean{proboszcz}{parish priest} and \mean{chłopak}{boy}. The former fulfills all of the criteria above, besides being perceived as at least 60\% masculine in the norming study, since it is not included there. However, it can be included in the experiment because it is a noun commonly used in Polish and, as a subtype of a priest, it counts as being perceived as masculine.

As \cite{skowronska_konstrukcje_2011} reports, the latter of the two additional nouns is problematic for Polish speakers. In a section dedicated specifically to the lexeme `chłopak' she reports that some Polish speakers are taught that the non-depreciative form of the noun is incorrect. Because of this, hybrid agreement is very commonly used with this particular noun. The author also looked through the National Corpus of Polish to see a few examples of the noun with attributive and predicate agreement and she finds that predicates with this noun are mostly seen with semantic agreement, while attributes occur almost exclusively with syntactic agreement. This is in line with Corbett's predictions and the current study can show whether this tendency extends to the further points of the Agreement Hierarchy.

All of the nouns chosen for this experiment can be found in the appendix. In (\ref{ex:depr-item}) there is an example of an experimental item with one of the nouns in depreciative form. 

\begin{exe}
  \ex\label{ex:depr-item} Example of an item in the experiment with depreciative forms in all conditions and with its comprehension question.
  \begin{xlist}
    \ex f/m1, prenominal
    \gll Na nagrania do Ameryki Południowej poleciały/polecieli reżysery.\\
    for shooting in America South flew:\textsc{f}/flew:\textsc{m1} directors\textsc{(m1)}\\
    \glt The directors flew to South America for shooting.\\
    
    \ex f/m1, postnominal
    \gll Reżysery poleciały/polecieli na nagrania do Ameryki Południowej.\\
    directors\textsc{(m1)} flew:\textsc{f}/flew:\textsc{m1} for shooting to America South\\
    \glt The directors flew to South America for shooting.\\
    
    \ex f/m1, relative pronoun
    \gll Reżysery, które/którzy lecą na nagrania do Ameryki Południowej, nie odpowiadają teraz na maile.\\
    directors\textsc{(m1)} who:\textsc{f}/who:\textsc{m1} are.flying for shooting to America South \textsc{NEG} responding now to emails\\
    \glt Directors, who are flying to South America to shoot, are not responding to emails right now.\\

    \ex comprehension question\\
    Czy w zdaniu wspomniano o Azji?\\
    \textsl{Did the sentence mention Asia?}
  \end{xlist}
\end{exe}

\subsubsection{Experimental items with profession nouns}
The process of selecting nouns for this experiment was similar to that for the corpus study. The study by \cite{szpyra-kozlowska_premiera_2019} was used to find masculine nouns, for which it is difficult to propose a feminine version. Nouns were arranged from the most times participants of the study refused to give a feminine version of the noun, to the least. This formed a ranking, from which the following nouns were excluded:
\begin{itemize}
\item \mean{ksiądz}{priest}, \mean{pastor}{pastor} and \mean{proboszcz}{parish priest}, because in Poland the population of female priests is almost nonexistent
\item offensive words (e.g. \mean{dureń}{fool})
\item words with less than a 1000 occurrences in the balanced version of the National Corpus of Polish (with prose excluded to make sure the words are used in everyday language)
\end{itemize}

Here each experimental item also required a sentence of context preceding the target sentence. The nouns used in this experiments are masculine and can refer to both men and women, so context is needed for participants to know that in this situation the noun refers to a woman. All of the items for this experiment can be found in the appendix and an example of an item can be found in (\ref{ex:prof-item}).

\begin{exe}
  \ex\label{ex:prof-item} Example of an item in the experiment with profession nouns in all conditions, with its context and the comprehension question.
  \begin{xlist}
    \ex context\\
    W tym tygodniu odwiedzają nas mamy uczniów z naszej klasy, żeby opowiedzieć o swojej pracy.\\
    \textsl{Moms of kids from our class are visiting us at school this week to talk about their careers.}
    
    \ex f/m1, prenominal
    \gll We wtorek przyszła/przyszedł architekt.\\
    on Tuesday came:\textsc{f}/came:\textsc{m1} architect\textsc{(m1)}\\
    \glt The architect came on Tuesday.\\
    
    \ex f/m1, postnominal
    \gll Architekt przyszła/przyszedł we wtorek.\\
    architect\textsc{(m1)} came:\textsc{f}/came:\textsc{m1} on Tuesday\\
    \glt The architect came on Tuesday.\\
    
    \ex f/m1, relative pronoun
    \gll Architekt, która/który przyjdzie we wtorek, należy do Rady Rodziców.\\
    architect\textsc{(m1)} who:\textsc{f}/who:\textsc{m1} will.come on Tuesday belongs to council of.parents\\
    \glt The architect who will come on Tuesday is a member of the Parents' Council.\\

    \ex comprehension question\\
    Czy dzień tygodnia, który był wspomniany w zdaniu, to czwartek?\\
    \textsl{Was Thursday the day of the week mentioned in the sentence?}
  \end{xlist}
\end{exe}

\subsubsection{Other items}
Both experiments also had 12 control items, half correct, to establish a threshold of acceptable and unacceptable sentences. The incorrect sentences contained an attraction error with the subject, the correct ones were the same but without the error. Half of the controls had human subjects and half had non-human, this was distributed evenly within the conditions. In the profession nouns experiment these items were preceded by a sentence of context, so that the experimental item would not be too easy to distinguish. An example of a control item can be found in (\ref{ex:control-item}).

\begin{exe}
  \ex\label{ex:control-item} Example of a control item in both conditions and the context sentence.
  \begin{xlist}
    \ex context\\
    Nie wszystkie dzieci chciałyby mieć wieczne wakacje.\\
    \textsl{Not all kids want to be on holidays forever.}

    \ex correct/incorrect
    \gll Syn sąsiadki jest szczęśliwy/szczęśliwa że idzie do szkoły.\\
    son\textsc{(m1)} of.neighbor\textsc{(f)} is happy:\textsc{m1}/happy:\textsc{f} that goes to school\\
    \glt The son of the neighbor is happy to go to school.\\
  \end{xlist}
\end{exe}

To make sure the main focus of the study is not too obvious to participants, there were also 24 filler items. As with control items there was a context sentence preceding the fillers in the profession noun experiment. An example can be found in (\ref{ex:filler-item}).

\begin{exe}
  \ex\label{ex:filler-item} Example of a filler item with its context.
  \begin{xlist}
    \ex context\\
    Szukam zeszytu w którym babcia zapisywała hasła.\\
    \textsl{I'm looking for the notebook where grandma wrote down her passcodes.}

    \ex 
    \gll Znalazłem na strychu sejf i nie wiem jak go otworzyć.\\
    I.found on attic strongbox and \textsc{neg} know how it open\\
    \glt I found a strongbox in the attic and I don't know how to open it.\\
  \end{xlist}
\end{exe}

\subsection{Participants}
Participants for both studies were recruited using Prolific. The target population on the platform was filtered to only include people who currently live in Poland, have Polish nationality and speak Polish as their first language. There were in total 60 participants (44 men and 16 women), their age ranged from 18 to 55. They were paid 3.20 GBP (3.7 EUR) for participation. This is 15.69 PLN for 20 minutes which corresponds to 47.07 PLN hourly, so around 150\% of minimum hourly wage in Poland as of 2026. 

\subsection{Hypotheses}
Based on the corpus study, one hypothesis is that syntactic agreement is preferred over semantic, so items with semantic agreement should have an overall lower acceptability rating than the items with syntactic agreement. The semantic agreement should, however, be rated higher than the ungrammatical control sentences.

Additionally, according to the stronger constraint of Corbett's Agreement Hierarchy, the acceptability of semantic agreement should increase as we go rightwards along the hierarchy. Here this means that semantic predicate agreement has a lower average acceptability rating than semantic relative pronoun agreement. Another hypothesis concerns the Hierarchy of Agreement Sources, based on which semantic agreement with split hybrids is less acceptable than with lexical hybrids. For this study the split hybrids are depreciative forms and lexical hybrids are profession nouns, therefore semantic agreement in the the profession nouns experiment should have an overall higher acceptability than semantic agreement in the depreciatives experiment. Finally, according to Corbett the target of agreement preceding the controller decreases the likelihood of semantic agreement, so in the experiments the prenominal condition should have a lower acceptability than the postnomial condition.

\section{Results of Experiment 1}

\section{Results of Experiment 2}
