---
title: "Full corpus results"
author: "Magdalena Borysiak"
date: "`r Sys.Date()`"
output: html_document
---

# Loading data

Importing necessary libraries:
```{r}
library(lme4)
library(dplyr)
library(ggplot2)
library(wesanderson)
library(stringr)
library(patchwork)
library(xtable)
library(data.table)
library(modeest)
library(ordinal)
```

Loading data:
```{r}
numbers <- sprintf("%03d", 0:190)
filenames <- paste0("NKJP_300M_", numbers, ".csv")

filepaths <- file.path("../agreement-occurrences", filenames)
agrs <- lapply(filepaths, read.csv) %>% bind_rows()

filepaths <- file.path("../agreement-occurrences/new-profession-nouns", filenames)
profs <- lapply(filepaths, read.csv) %>% bind_rows()
rm(list=setdiff(ls(), c("agrs", "profs")))
```
`agrs` is the dataframe with all data that I extracted for my M1 and `profs` is the dataframe with data extracted when I wrote scripts to also extract information about appositions for profession nouns.

I got some terminology mixed up when I was first preparing this dataset and I'm afraid changing it now will result in issues with code, so:
- in this document words *source* and *controller* are used interchangeably, even though in the literature they mean different things
- one type of agreement is called *predicative*, when I think it should be called *predicate*

Up to here was saved as loaded.RData

# Preparing data
I will merge the dataframes for a moment to prepare all of the data for further work, then I will separate the profession nouns to do more specific work on them.
```{r}
agrs %>% filter(source_cat != "profession") -> agrs
agrs <- bind_rows(agrs, profs)
rm(profs)
```

Looking at the summary to change the types of variables:
```{r}
summary(agrs)
```

# Exploring, filtering
Before I look at the summary again I know I should filter out one lexeme.
```{r}
agrs %>% filter(source_lexeme == "pasztet") %>% select(sentence)
```
Majority of these occurrences are not in the sense that was intended for the study, so I am excluding these.
```{r}
agrs %>% filter(source_lexeme != "pasztet") -> agrs
```

Looking again at the summary of data, based on this I will filter some things out.
```{r}
summary(agrs)
```

## Large distances
The maximum distance can be large, I will look at the values from the highest.
```{r}
agrs %>% group_by(distance) %>% count() %>% arrange(desc(distance))
```
Large distances are most likely parsing mistakes, so I will remove them.
```{r}
c(mean(agrs$distance) - 2*sd(agrs$distance), mean(agrs$distance) + 2*sd(agrs$distance))

agrs %>% 
  filter(distance > mean(distance) - 2*sd(distance)) %>% 
  filter(distance < mean(distance) + 2*sd(distance)) -> agrs
```

## Relative pronouns preceding the controller
The relative pronoun shouldn't precede the controller of agreement, so I'm also removing the occurrences where this happened.
```{r}
agrs %>% 
  filter(!(target_first == 1 & type_of_agreement == "relpron")) -> agrs
```

## Targets with digits
Removing instances of digits agreeing with words:
```{r}
agrs %>% 
  filter(is.na(as.numeric(target_form))) -> agrs
```

## Acronyms in depreciatives
There are some acronyms which were tagged as depreciative forms.
```{r}
agrs %>% 
  filter(source_cat == "depreciative") %>% 
  filter(str_detect(source_lexeme, '[:upper:]') & !str_detect(source_lexeme, '[:lower:]')) %>% 
  group_by(source_lexeme) %>% 
  count()
```
Acronyms are outside the scope of this study, so I will remove these.
```{r}
agrs %>% 
  filter(source_cat != "depreciative" | !(str_detect(source_lexeme, '[:upper:]') & !str_detect(source_lexeme, '[:lower:]'))) -> agrs
```

## Potential coordinations
When I'm not looking at agreement with depreciative forms, target or controller shouldn't be plural because then they might be a part of a coordination (which I'm not considering here).
```{r}
agrs %>% 
  filter((!grepl("pl", source_morph_descr)&!grepl("pl", target_morph_descr))|source_cat == "depreciative") -> agrs
```

## Lists in sentences
I know that in the corpus sometimes there are sentences that are actually just lists of people with certain functions, which gives me a lot of results that aren't actually useful.
```{r}
agrs %>% group_by(sentence_id) %>% count() %>% arrange(desc(n))
```
I looked into the first two sentences here (`7708943` and `2517756`) and I found these are rather lists than regular sentences, so here is a list of sentences I found this way and I will remove them.
```{r}
agrs %>% 
  filter(sentence_id %in% c(7708943, 7708935, 7708932, 2517756, 7708936, 7708941)) %>% 
  group_by(sentence, sentence_id) %>% 
  count()
```
Removing sentences which are not actually sentences:
```{r}
agrs %>% 
  filter(!(sentence_id %in% c(7708943, 7708935, 7708932, 2517756, 7708936, 7708941))) -> agrs
```

## Non-m1 nouns
Sometimes the grammatical gender of the controller was marked not as m1. This is expected for depreciative forms (the tagging system assumes m2 instead of m1 for them) and for some gendered nouns, but I want to make sure that it's correct in those categories and filter out the mistakes in the profession noun category.
```{r}
agrs %>% 
  filter(source_gram_gender != "m1") %>% 
  group_by(source_cat) %>% 
  count()

agrs %>% 
  filter(source_cat=="depreciative") %>% 
  group_by(source_gram_gender) %>% 
  count()

agrs %>% 
  filter(source_cat=="gendered") %>% 
  group_by(source_lexeme, source_gram_gender) %>% 
  count()
```
Depreciative forms are all m2, so I will just keep that in mind for when I mutate the variable for semantic/syntactic agreement.

### Chłopaczysko
In gendered nouns there is one that requires revision: "chłopaczysko" can be either m1 or n according to the dictionary, so I need to judge if these occurrences are n and can be kept or m1 and should be discarded.
```{r}
agrs %>% 
  filter(source_lexeme == "chłopaczysko") %>% 
  select(sentence, source_morph_descr)
```
All of the occurrences tagged as m1 are in nominative or vocative, for which there is no distinct m1 form, so I assume these are n.
```{r}
agrs %>% 
  mutate(source_gram_gender = if_else(source_lexeme=="chłopaczysko", "n", source_gram_gender)) -> agrs
```

Now I will look into the profession nouns, because 1297 of them had non-m1 gender, when it should be m1 for all of them. I want a list of all nouns that were marked as non-m1 and how many occurrences were there of each one.
```{r}
agrs %>% 
  filter(source_cat == "profession") %>% 
  filter(source_gram_gender != "m1") %>% 
  group_by(source_lexeme) %>% 
  count() %>% 
  arrange(desc(n))
```

### Zbieg
```{r}
agrs %>% 
  filter(source_cat == "profession") %>% 
  filter(source_gram_gender != "m1") %>% 
  filter(source_lexeme == "zbieg") %>% 
  select(sentence, source_form, target_form)
```
The occurrences for the "zbieg" lexeme are probably not a parsing mistake; this word can be also used in the colocation "zbieg okoliczności" meaning "coincidence". 
```{r}
agrs %>% 
  filter(source_cat == "profession" & source_lexeme == "zbieg") %>% 
  group_by(source_gram_gender) %>% 
  count()
```
I will assume all m3 occurrences talk about a coincidence and leave in the dataset only the m1 occurrences.
```{r}
agrs %>% 
  filter(!(source_lexeme == "zbieg" & source_gram_gender == "m3")) -> agrs
```

### Upiór
```{r}
agrs %>% 
  filter(source_cat == "profession") %>% 
  filter(source_gram_gender != "m1") %>% 
  filter(source_lexeme == "upiór") %>% 
  select(sentence, source_form, target_form)
```
This word can be also used to mean a haunting memory, so this is a more metaphorical meaning. I will filter out these occurrences.
```{r}
agrs %>% 
  filter(!(source_lexeme == "upiór" & source_gram_gender != "m1")) -> agrs
```

### Subiekt
```{r}
agrs %>% 
  filter(source_gram_gender != "m1") %>% 
  filter(source_lexeme == "subiekt") %>% 
  select(sentence, source_form, target_form)
```
These seem to be a parsing mistake, those sentences mostly mention a person referred to as "subiekt". I will recode these as m1.
```{r}
agrs %>% 
  mutate(source_gram_gender = if_else(source_lexeme=="subiekt", "m1", source_gram_gender)) -> agrs
```

### Lump
```{r}
agrs %>% 
  filter(source_gram_gender != "m1") %>% 
  filter(source_lexeme == "lump") %>% 
  select(sentence, sentence_id, source_form, target_form, file_id)
```
I can't say with certainty that all of these refer to a person (and not to "old clothes", which is also possible with this lexeme), so I will filter these out.
```{r}
agrs %>% 
  filter(!(source_lexeme == "lump" & source_gram_gender != "m1")) -> agrs
```

### Dureń
```{r}
agrs %>% 
  filter(source_gram_gender != "m1") %>% 
  filter(source_lexeme == "dureń") %>% 
  select(sentence, source_form, target_form, sentence_id, file_id)
```
All of these seem to be referring to a person, therefore I will recode them to m1.
```{r}
agrs %>% 
  mutate(source_gram_gender = if_else(source_lexeme=="dureń", "m1", source_gram_gender)) -> agrs
```

### Pajac
```{r}
agrs %>% 
  filter(source_gram_gender != "m1") %>% 
  filter(source_lexeme == "pajac") %>% 
  select(sentence, sentence_id, source_form, target_form)
```
Only of 3 these 10 clearly refer to a person, so these will be recoded to m1 and the rest will be removed.
```{r}
agrs %>% 
  filter(source_lexeme != "pajac" | (source_gram_gender == "m1" | sentence_id == 15054693)) -> agrs

agrs %>% 
  mutate(source_gram_gender = if_else(source_lexeme=="pajac", "m1", source_gram_gender)) -> agrs
```

### Nurek
```{r}
agrs %>% 
  filter(source_gram_gender != "m1") %>% 
  filter(source_lexeme == "nurek") %>% 
  select(sentence, sentence_id, source_form, target_form)
```
Both of those occurrences have the noun refer to the activity of diving and not a person, so I will filter them out.
```{r}
agrs %>% 
  filter(!(source_lexeme == "nurek" & source_gram_gender != "m1")) -> agrs
```

### Finishing
```{r}
agrs %>% 
  filter(source_gram_gender != "m1") %>% 
  group_by(source_cat) %>% 
  count()
```
All of these should be non-m1, so I can move on.

## Target gender
With depreciative forms I expect to see any target gender, with gendered nouns the target gender depends on the noun and with profession nouns the target gender should always be m1 or f.
```{r}
agrs %>% 
  filter(source_cat == "depreciative") %>% 
  group_by(source_gram_gender, source_other_gender, target_gender) %>% 
  count()

agrs %>% 
  filter(source_cat == "gendered") %>% 
  group_by(source_gram_gender, source_other_gender, target_gender) %>% 
  count()

agrs %>% 
  filter(source_cat == "profession") %>% 
  group_by(source_gram_gender, source_other_gender, target_gender) %>% 
  count()
```
As for profession nouns, there are too many to look through manually, therefore I am removing from the dataset all that do not have one of the genders in the controller.
```{r}
agrs %>% 
  filter(source_cat != "profession" | !(target_gender != source_gram_gender & target_gender != source_other_gender)) -> agrs
```

I want to look into the situations where the target gender with gendered is neither the grammatical not the other gender of the source.
```{r}
agrs %>% 
  filter(source_cat == "gendered") %>% 
  filter(target_gender != source_gram_gender & target_gender != source_other_gender) %>% 
  select(source_form, source_gram_gender, source_other_gender, target_form, target_gender, sentence, sentence_id)
```
After manually reviewing these I found parsing mistakes and I will remove a few occurrences and recode one.
```{r}
agrs %>% 
  filter(!((sentence_id == 376931 & target_form == "tym")|
         (sentence_id == 8977510 & target_form == "egzotycznego")|
         (sentence_id == 11585730 & target_form == "super")|
         (sentence_id == 15147190 & target_form == "sam")|
         (sentence_id == 15267353 & target_form == "wniebogłosy")|
         (sentence_id == 16638227 & target_form == "co"))) -> agrs

agrs %>% 
  mutate(target_gender = if_else(sentence_id == 778950 & target_form == "którego",
                                 "n", target_gender)) -> agrs
```

Making sure that the target gender is cleared up now.
```{r}
agrs %>% 
  filter(source_cat == "depreciative") %>% 
  group_by(source_gram_gender, source_other_gender, target_gender) %>% 
  count()

agrs %>% 
  filter(source_cat == "gendered") %>% 
  group_by(source_gram_gender, source_other_gender, target_gender) %>% 
  count()

agrs %>% 
  filter(source_cat == "profession") %>% 
  group_by(source_gram_gender, source_other_gender, target_gender) %>% 
  count()
```

Finally, I will merge all masculine genders into m, because the distinction between them is not useful for the analysis from now on.
```{r}
agrs %>% 
  mutate(source_gram_gender = case_when(source_cat == "depreciative" ~ "f",
                                        grepl("m", source_gram_gender) ~ "m",
                                        T ~ source_gram_gender),
         source_other_gender = case_when(grepl("m", source_other_gender) ~ "m",
                                         T ~ source_other_gender),
         target_gender = case_when(source_cat == "depreciative"
                                   & target_gender == "m1" ~ "m",
                                   source_cat == "depreciative"
                                   & target_gender != "m" ~ "f",
                                   grepl("m", target_gender) ~ "m",
                                   T ~ target_gender)) -> agrs
```

Up to here was saved as cleaned.RData

# Double agreements in profession nouns
*SKIP THIS SECTION FOR NOW*

In this study I look at attributive, predicate and relative pronoun agreements. In this part I want to see which genders are used when there are two types of agreement with one controller. With the example of attributive and predicate agreement, I want to know which one of the following is preferred in the corpus.
- Af N Vf
- Am N Vf
- Af N Vm
- Am N Vm
More specifically I want to know if the first agreement being semantic makes it more likely that the second will be semantic too.

In the following subsections I will look for double attributive/predicate, attributive/relative pronoun and predicate/relative pronoun agreements with profession nouns only. I will not include other nouns for this part, because there is a lot less of them and it would introduce another variable that can't be explored carefully enough.
```{r}
agrs %>% filter(source_cat == "profession") -> profs
```

## Preparing tables
Chunks in this section are commented out because they generate a table that is saved locally and they take a lot of time.
```{r}
# profs %>%
#   group_by(sentence_id) %>%
#   count() %>%
#   filter(n == 2) %>%
#   select(sentence_id) -> multiples
# multiples <- multiples$sentence_id
```
Iterating over this list to make a dataset.
```{r}
# AVdoubles <- data.frame()
# for (id in multiples) {
#   profs %>% filter(sentence_id == id) -> current_sentence
#   if (current_sentence[1,]$type_of_agreement == "predicative"
#       & current_sentence[2,]$type_of_agreement == "attributive"
#       & current_sentence[1,]$source_lexeme == current_sentence[2,]$source_lexeme) {
#     cbind(current_sentence[1,], current_sentence[2,]) -> newrow
#   }
#   if (current_sentence[2,]$type_of_agreement == "predicative"
#       & current_sentence[1,]$type_of_agreement == "attributive"
#       & current_sentence[1,]$source_lexeme == current_sentence[2,]$source_lexeme) {
#     cbind(current_sentence[1,], current_sentence[2,]) -> newrow
#   }
#   AVdoubles <- rbind(AVdoubles, newrow)
# }
```
Some rows are duplicated so I will remove them.
```{r}
# AVdoubles[!duplicated(AVdoubles),] -> AVdoubles
```

Saving this because (1) it took a long time to generate this dataset and (2) reading this table from a file fixes the issue with repeating column names.
```{r}
# write.csv(AVdoubles, "AV-doubles.csv", row.names = F)
```

Removing some useless variables from the environment.
```{r}
# rm(current_sentence, newrow, id, AVdoubles)
```

Now I will do the same for the attributive/relative pronoun agreements.
```{r}
# ARPdoubles <- data.frame()
# for (id in multiples) {
#   prof %>% filter(sentence_id == id) -> current_sentence
#   if (current_sentence[1,]$type_of_agreement == "relpron"
#       & current_sentence[2,]$type_of_agreement == "attributive"
#       & current_sentence[1,]$source_lexeme == current_sentence[2,]$source_lexeme) {
#     cbind(current_sentence[1,], current_sentence[2,]) -> newrow
#     ARPdoubles <- rbind(ARPdoubles, newrow)
#   }
#   if (current_sentence[2,]$type_of_agreement == "relpron"
#       & current_sentence[1,]$type_of_agreement == "attributive"
#       & current_sentence[1,]$source_lexeme == current_sentence[2,]$source_lexeme) {
#     cbind(current_sentence[1,], current_sentence[2,]) -> newrow
#     ARPdoubles <- rbind(ARPdoubles, newrow)
#   }
# }
#
# ARPdoubles[!duplicated(ARPdoubles),] -> ARPdoubles
#
# write.csv(ARPdoubles, "ARP-doubles.csv", row.names = F)
#
# rm(current_sentence, newrow, id, ARPdoubles)
```

And for predicate/relative pronoun agreements.
```{r}
# VRPdoubles <- data.frame()
# for (id in multiples) {
#   prof %>% filter(sentence_id == id) -> current_sentence
#   if (current_sentence[1,]$type_of_agreement == "relpron"
#       & current_sentence[2,]$type_of_agreement == "predicative"
#       & current_sentence[1,]$source_lexeme == current_sentence[2,]$source_lexeme) {
#     cbind(current_sentence[1,], current_sentence[2,]) -> newrow
#     VRPdoubles <- rbind(VRPdoubles, newrow)
#   }
#   if (current_sentence[2,]$type_of_agreement == "relpron"
#       & current_sentence[1,]$type_of_agreement == "predicative"
#       & current_sentence[1,]$source_lexeme == current_sentence[2,]$source_lexeme) {
#     cbind(current_sentence[1,], current_sentence[2,]) -> newrow
#     VRPdoubles <- rbind(VRPdoubles, newrow)
#   }
# }
#
# VRPdoubles[!duplicated(VRPdoubles),] -> VRPdoubles
#
# write.csv(VRPdoubles, "VRP-doubles.csv", row.names = F)
# 
# rm(current_sentence, newrow, id, VRPdoubles, multiples)
```

## Attributive and predicate
Reading the table from file.
```{r}
read.csv("AV-doubles.csv") -> AVdoubles
```

I'm mutating a column to see all of the combinations I'm interested in: AfVf, AmVf, AfVm, AmVm. Later grouping will be based on this one instead of grouping based on three different variables.
```{r}
AVdoubles %>% 
  mutate(condition = case_when(type_of_agreement == "attributive"
                               & target_gender == "f"
                               & type_of_agreement.1 == "predicative"
                               & target_gender.1 == "f" 
                               ~ "AfVf",
                               type_of_agreement == "attributive"
                               & target_gender == "m"
                               & type_of_agreement.1 == "predicative"
                               & target_gender.1 == "f" 
                               ~ "AmVf",
                               type_of_agreement == "attributive"
                               & target_gender == "f"
                               & type_of_agreement.1 == "predicative"
                               & target_gender.1 == "m" 
                               ~ "AfVm",
                               type_of_agreement == "attributive"
                               & target_gender == "m"
                               & type_of_agreement.1 == "predicative"
                               & target_gender.1 == "m" 
                               ~ "AmVm",
                               type_of_agreement.1 == "attributive"
                               & target_gender.1 == "f"
                               & type_of_agreement == "predicative"
                               & target_gender == "f" 
                               ~ "AfVf",
                               type_of_agreement.1 == "attributive"
                               & target_gender.1 == "m"
                               & type_of_agreement == "predicative"
                               & target_gender == "f" 
                               ~ "AmVf",
                               type_of_agreement.1 == "attributive" 
                               & target_gender.1 == "f" 
                               & type_of_agreement == "predicative" 
                               & target_gender == "m" 
                               ~ "AfVm",
                               type_of_agreement.1 == "attributive" 
                               & target_gender.1 == "m" 
                               & type_of_agreement == "predicative" 
                               & target_gender == "m" 
                               ~ "AmVm")) -> AVdoubles
```

```{r}
AVdoubles %>% 
  filter(target_gender != "n" & target_gender.1 != "n") %>% 
  mutate(target_gender = ifelse(target_gender != "f", "m", "f"),
         target_gender.1 = ifelse(target_gender.1 != "f", "m", "f")) %>%
  mutate(condition = case_when(type_of_agreement == "attributive"
                               & target_gender == "f"
                               & type_of_agreement.1 == "predicative"
                               & target_gender.1 == "f" 
                               ~ "AfVf",
                               type_of_agreement == "attributive"
                               & target_gender == "m"
                               & type_of_agreement.1 == "predicative"
                               & target_gender.1 == "f" 
                               ~ "AmVf",
                               type_of_agreement == "attributive"
                               & target_gender == "f"
                               & type_of_agreement.1 == "predicative"
                               & target_gender.1 == "m" 
                               ~ "AfVm",
                               type_of_agreement == "attributive"
                               & target_gender == "m"
                               & type_of_agreement.1 == "predicative"
                               & target_gender.1 == "m" 
                               ~ "AmVm",
                               type_of_agreement.1 == "attributive"
                               & target_gender.1 == "f"
                               & type_of_agreement == "predicative"
                               & target_gender == "f" 
                               ~ "AfVf",
                               type_of_agreement.1 == "attributive"
                               & target_gender.1 == "m"
                               & type_of_agreement == "predicative"
                               & target_gender == "f" 
                               ~ "AmVf",
                               type_of_agreement.1 == "attributive" 
                               & target_gender.1 == "f" 
                               & type_of_agreement == "predicative" 
                               & target_gender == "m" 
                               ~ "AfVm",
                               type_of_agreement.1 == "attributive" 
                               & target_gender.1 == "m" 
                               & type_of_agreement == "predicative" 
                               & target_gender == "m" 
                               ~ "AmVm")) -> AVdoubles
```


Getting counts for each Af/m and Vf/m combination, also grouping by ordering and showing the mean distance between each target and controller in each group. The AmVm condition is not actually relevant for me because if there is no feminine agreement then it's likely that the referent was male.
```{r}
AVdoubles %>% 
  filter(condition != "AmVm") %>% 
  group_by(condition, target_first, target_first.1) %>% 
  summarise(n = n(), distance = mean(distance), distance.1 = mean(distance.1))
```

```{r}
AVdoubles %>% 
  filter(condition == "AfVf" & target_first == 0 & target_first.1 == 0) %>% 
  select(target_form, target_form.1, sentence)
```







# Filtering profession nouns
Here I will filter the profession noun occurrences based on their appositions, so that I can add them to the dataset with gendered nouns and depreciative forms and compute models.

I need to mark missing appositions as NAs.
```{r}
agrs %>% filter(source_cat == "profession") -> profs

profs[profs==""] <- NA
```

I want to look at occurrences of plural appositions with singular controllers.
```{r}
profs %>% 
  filter(grepl("pl", appos_morph_descr)) %>% 
  select(source_form, target_form, apposition, sentence)
```
There are only 21 occurrences and they are either parsing mistakes or they signal that the controller was a part of coordination and the apposition actually refers to the whole coordination and not just the controller. Because of that the apposition is useless in this case and I will remove those occurrences.
```{r}
profs %>% 
  filter(!grepl("pl", appos_morph_descr)) -> profs
```

I need to know the gender of the appositions.
```{r}
profs %>% 
  filter(!is.na(apposition)) %>% 
  mutate(appos_gender = case_when(grepl(":m1", appos_morph_descr) ~ "m1",
                                  grepl(":m2", appos_morph_descr) ~ "m2",
                                  grepl(":m3", appos_morph_descr) ~ "m3",
                                  grepl(":n", appos_morph_descr) ~ "n",
                                  grepl(":f", appos_morph_descr) ~ "f")) %>% 
  filter(!is.na(appos_gender)) -> profs
```

I want to see how many occurrences were there with different genders of appositions.
```{r}
profs %>%
  group_by(appos_gender) %>% 
  count()
```

I want to use only the occurrences with feminine apositions.
```{r}
profs %>% 
  filter(appos_gender == "f") %>% 
  select(!appos_gender) -> profs
```

I also want to have a variable with the ordering of controller, target and apposition, so I have a function that detects that ordering in the occurrence (ideally it would have been done at the stage of extracting occurrences from dependency trees, but I didn't think about this at the time).
```{r}
ordering <- function(sentence, target, source, apposition, target_first, distance) {
  if(str_count(sentence, target) > 1 
     | str_count(sentence, source) > 1 
     | str_count(sentence, apposition) > 1) return (NA)
  
  t <- str_locate(sentence, target)
  s <- str_locate(sentence, source)
  if (target_first == 0) {
    as <- str_count(substr(sentence, 0, s[1]), apposition)
    if(as == 1) return("AST")
    if(as > 1) return(NA)
    if(as == 0) {
      at <- str_count(substr(sentence, t[2], 2000), apposition)
      if(at == 1) return("STA")
      if(at > 1) return(NA)
      if(at == 0 & distance != 0) return("SAT") else return(NA)
    }
  } else {
    as <- str_count(substr(sentence, 0, t[1]), apposition)
    # print(as)
    # print(sentence)
    if(as == 1) return("ATS")
    if(as > 1) return(NA)
    if(as == 0) {
      at <- str_count(substr(sentence, s[2], 2000), apposition)
      if(at == 1) return("TSA")
      if(at > 1) return(NA)
      if(at == 0 & distance != 0) return("TAS") else return(NA)
    }
  }
}
```

Here I apply the function to my data.
```{r}
new_col <- character(nrow(profs))
for (i in 1:nrow(profs)) {
  ordering(str_to_lower(profs[i,]$sentence), 
           profs[i,]$target_form, 
           profs[i,]$source_form, 
           profs[i,]$apposition, 
           profs[i,]$target_first, 
           profs[i,]$distance) -> new_row
  new_col[i] <- new_row
}
profs %>% 
  mutate(order = new_col) -> profs
rm(i, new_col, new_row)
```

I want to see if the function managed to assign an order to every occurrence.
```{r}
profs %>% filter(is.na(order))
```

If I have time I will look through them and assign the order manually but for now I will filter these out.
```{r}
profs %>% filter(!is.na(order)) -> profs
```

Finally I can merge these verified profession nouns occurrences with the rest of the dataset.
```{r}
agrs %>% 
  filter(source_cat != "profession") -> agrs

agrs <- bind_rows(agrs, profs)
rm(list=setdiff(ls(), "agrs"))
```

Up to here was saved as filtered-prof.RData

# Descriptive stats
Adjusting the types:
```{r}
agrs %>% 
  mutate(type_of_text = as.factor(type_of_text),
         source_cat = as.factor(source_cat),
         source_gram_gender = as.factor(source_gram_gender),
         source_other_gender = as.factor(source_other_gender),
         target_gender = as.factor(target_gender),
         type_of_agreement = as.factor(type_of_agreement),
         target_first = as.factor(target_first)) -> agrs
```

A table with values of possible genders for all categories of controllers.
```{r}
agrs %>% 
  group_by(source_cat, source_gram_gender, source_other_gender, target_gender) %>% 
  count()
```
For depreciative forms:
- semantic agreement is when target=other (profesory przyszli)
- syntactic agreement is when target=gram (profesory przyszły)

For profession nouns:
- semantic agreement is when target=other (wójt przyszła)
- syntactic agreement is when target=gram (wójt przyszedł)

For gendered nouns:
- semantic agreement is when target=other (dziewczę przyszła)
- syntactic agreement is when target=gram (dziewczę przyszło)

```{r}
# first should be semantic
agrs %>% 
  filter(source_cat == "gendered") %>% 
  filter(target_gender != source_gram_gender) %>% 
  select(source_form, target_form)

# second should be syntactic
agrs %>% 
  filter(source_cat == "gendered") %>% 
  filter(target_gender == source_gram_gender) %>% 
  select(source_form, target_form)
```

Mutating a boolean column for semantic agreement:
```{r}
agrs %>%
  mutate(semantic_agr = as.factor(case_when(target_gender != source_gram_gender ~ 1,
                                            target_gender == source_gram_gender ~ 0))) -> agrs
```

```{r}
agrs %>% 
  group_by(semantic_agr) %>% 
  count()
```

Some tables as an overview:
```{r}
# profession nouns
agrs %>% 
  mutate(semantic_agr = as.numeric(levels(semantic_agr))[semantic_agr],
         type_of_agreement = as.character(type_of_agreement)) %>% 
  filter(source_cat == "profession") %>% 
  group_by(type_of_agreement) %>% 
  summarise(n = n(),
            sem_n = sum(semantic_agr),
            syn_n = n-sem_n) %>% 
  bind_rows(summarise_all(., ~if(is.numeric(.)) sum(.) else "")) %>% 
  mutate(sem_prop = paste0(round(sem_n/n*100, 2), "%"),
         syn_prop = paste0(round(syn_n/n*100, 2), "%")) %>% 
  relocate(sem_prop, .after = sem_n) #%>% xtable()

# gendered nouns
agrs %>% 
  mutate(semantic_agr = as.numeric(levels(semantic_agr))[semantic_agr],
         type_of_agreement = as.character(type_of_agreement)) %>% 
  filter(source_cat == "gendered") %>% 
  group_by(type_of_agreement) %>% 
  summarise(n = n(),
            sem_n = sum(semantic_agr),
            syn_n = n-sem_n) %>% 
  bind_rows(summarise_all(., ~if(is.numeric(.)) sum(.) else "")) %>% 
  mutate(sem_prop = paste0(round(sem_n/n*100, 2), "%"),
         syn_prop = paste0(round(syn_n/n*100, 2), "%")) %>% 
  relocate(sem_prop, .after = sem_n) #%>% xtable()

# depreciative forms
agrs %>% 
  mutate(semantic_agr = as.numeric(levels(semantic_agr))[semantic_agr],
         type_of_agreement = as.character(type_of_agreement)) %>% 
  filter(source_cat == "depreciative") %>% 
  group_by(type_of_agreement) %>% 
  summarise(n = n(),
            sem_n = sum(semantic_agr),
            syn_n = n-sem_n) %>% 
  bind_rows(summarise_all(., ~if(is.numeric(.)) sum(.) else "")) %>% 
  mutate(sem_prop = paste0(round(sem_n/n*100, 2), "%"),
         syn_prop = paste0(round(syn_n/n*100, 2), "%")) %>% 
  relocate(sem_prop, .after = sem_n) #%>% xtable()

# all nouns
agrs %>% 
  mutate(semantic_agr = as.numeric(levels(semantic_agr))[semantic_agr],
         type_of_agreement = as.character(type_of_agreement)) %>% 
  group_by(type_of_agreement) %>% 
  summarise(n = n(),
            sem_n = sum(semantic_agr),
            syn_n = n-sem_n) %>% 
  bind_rows(summarise_all(., ~if(is.numeric(.)) sum(.) else "")) %>% 
  mutate(sem_prop = paste0(round(sem_n/n*100, 2), "%"),
         syn_prop = paste0(round(syn_n/n*100, 2), "%")) %>% 
  relocate(sem_prop, .after = sem_n) #%>% xtable()
```


The hypotheses are the following:

1. There is more semantic agreement with relative pronouns than with predicates and more predicate semantic agreement than attributive.

2. The larger the distance between the target and the source the more likely it is there will be semantic agreement.

3. If the target is before the source it is less likely there will be semantic agreement.
   
   - reference level: 1, having 0 for target_first increases the probability of semantic agreement
   
4. Lexical hybrids have more semantic agreement than split hybrids.

   - reference level: depreciative forms

First look at how many semantic agreements are there with different values of relevant variables.
```{r}
table(agrs$type_of_agreement, agrs$semantic_agr)
table(agrs$source_cat, agrs$semantic_agr)
table(agrs$target_first, agrs$semantic_agr)
table(agrs$distance, agrs$semantic_agr)
table(agrs$order, agrs$semantic_agr)
```

Tables showing percentages of semantic agreement for each noun and target.
```{r}
agrs %>% 
  mutate(semantic_agr = as.numeric(levels(semantic_agr))[semantic_agr],
         type_of_agreement = as.character(type_of_agreement)) %>% 
  filter(source_cat == "profession") %>% 
  group_by(type_of_agreement, target_first) %>% 
  summarise(n = n(),
            sem_n = sum(semantic_agr)) %>% 
  mutate(sem_prop = paste0(n, " (", round(sem_n/n*100, 2), "%)")) %>% 
  select(type_of_agreement, target_first, sem_prop) %>% 
  tidyr::pivot_wider(names_from = target_first,
                     values_from = sem_prop) %>% xtable()

agrs %>% 
  mutate(semantic_agr = as.numeric(levels(semantic_agr))[semantic_agr],
         type_of_agreement = as.character(type_of_agreement)) %>% 
  filter(source_cat == "gendered") %>% 
  group_by(type_of_agreement, target_first) %>% 
  summarise(n = n(),
            sem_n = sum(semantic_agr)) %>% 
  mutate(sem_prop = paste0(n, " (", round(sem_n/n*100, 2), "%)")) %>% 
  select(type_of_agreement, target_first, sem_prop) %>% 
  tidyr::pivot_wider(names_from = target_first,
                     values_from = sem_prop) %>% xtable()

agrs %>% 
  mutate(semantic_agr = as.numeric(levels(semantic_agr))[semantic_agr],
         type_of_agreement = as.character(type_of_agreement)) %>% 
  filter(source_cat == "depreciative") %>% 
  group_by(type_of_agreement, target_first) %>% 
  summarise(n = n(),
            sem_n = sum(semantic_agr)) %>% 
  mutate(sem_prop = paste0(n, " (", round(sem_n/n*100, 2), "%)")) %>% 
  select(type_of_agreement, target_first, sem_prop) %>% 
  tidyr::pivot_wider(names_from = target_first,
                     values_from = sem_prop) %>% xtable()
```


plots!
```{r}
agrs %>% 
  ggplot(aes(x = type_of_agreement, fill = semantic_agr)) + 
  geom_bar(position = "fill") + 
  labs(x = "Target of agreement",
       y = "Proportion",
       title = "Proportion of syntactic and semantic agreements in types of targets of agreement") +
  scale_fill_manual(name = "Agreement",
                    labels = c("Syntactic", "Semantic"),
                    values = wes_palette(n=2, name="Darjeeling1")) + 
  scale_x_discrete(labels = c("Attribute", "Predicate", "Relative pronoun")) -> semXtargets
semXtargets

agrs %>% 
  ggplot(aes(x = distance, fill = semantic_agr)) + 
  geom_bar(position = "fill") + 
  labs(x = "Agreement",
       y = "Proportion",
       title = "Counts of syntactic and semantic agreements in different distances between target and source") +
  scale_fill_discrete(name = "Agreement",
                      labels = c("Syntactic", "Semantic"))

agrs %>% 
  ggplot(aes(x = target_first, fill = semantic_agr)) + 
  geom_bar(position = "fill") + 
  labs(x = "Target-source order",
       y = "Proportion",
       title = "Proportion of syntactic and semantic agreements possible target-source orderings") +
  scale_fill_discrete(name = "Agreement",
                      labels = c("Syntactic", "Semantic")) +
  scale_x_discrete(labels = as_labeller(c(`0` = "Controller first", 
                                          `1` = "Target first")))

agrs %>% 
  ggplot(aes(x = source_cat, fill = semantic_agr)) + 
  geom_bar(position = "fill") + 
  labs(x = "Type of noun",
       y = "Proportion",
       title = "Proportion of syntactic and semantic agreements in split and lexical hybrids") +
  scale_fill_discrete(name = "Agreement",
                      labels = c("Syntactic", "Semantic")) +
  scale_x_discrete(labels = as_labeller(c(`depreciative` = "Depreciative", 
                                          `gendered` = "Gendered",
                                          `profession` = "Profession")))

agrs %>% 
  filter(source_cat == "profession") %>% 
  mutate(order = str_replace(order, "S", "C")) %>% 
  ggplot(aes(x = order, fill = semantic_agr)) +
  geom_bar(position = "fill") + 
  labs(x = "Order of target-controller-apposition",
       y = "Proportion", 
       title = "Proportion of syntactic and semantic agreements with different orders") +
  scale_fill_manual(name = "Agreement",
                    labels = c("Syntactic", "Semantic"),
                    values = wes_palette(n=2, name="Darjeeling1")) -> semXorder
semXorder
```
The first plot references the first hypothesis, which is that the likelihood of semantic agreement follows the pattern attributive < predicate < relative pronoun. This tendency is not visible in the data, attributive and predicate agreement have almost the same proportion, but there is a jump in semantic agreement with relative pronouns.

The second plot references the second hypothesis, which is that likelihood of semantic agreement increases with distance between the target and controller of agreement. This is somewhat visible in the plot, but maybe this should be modeled to see what is the tendency actually, because there are only `r (agrs %>% filter(distance == 9) %>% count())$n` occurrences with distance 9 and `r (agrs %>% filter(distance == 11) %>% count())$n` with distance 11.

We also hypothesise that target before controller makes semantic agreement less likely and in the third plot there is a small difference in this direction.

The fourth hypothesis is that lexical hybrids (here gendered nouns and profession nouns) have more semantic agreement than split hybrids (here depreciative forms). The fourth plot shows that one of the kinds of lexical hybrids, gendered nouns, have very little semantic agreement, then the split hybrids have some more and the other kind of lexical hybrids, profession nouns, have majority semantic agreement.

There was no particular hypothesis about ordering of target-controller-apposition with profession nouns, but I suspected that apposition intervening between target and controller might make semantic agreement more likely. Here it looks like it might be about apposition appearing before controller. I think when this happens, speakers register the gender of the referent as the one from apposition and not from the hybrid noun and that's why semantic agreement is more likely there.

```{r}
png(file="../writing/M2/images/semXorder.png", 
    width = 20,
    height = 8,
    units = "cm",
    res = 500)

semXorder & labs(title = NULL)

dev.off()
```

Now I will filter out the profession nouns occurrences with orders where apposition appears before the controller and look at the same plots made from this subset of data.

```{r}
agrs %>% 
  filter(!(order %in% c("AST", "ATS", "TAS"))) %>% 
  ggplot(aes(x = type_of_agreement, fill = semantic_agr)) + 
  geom_bar(position = "fill") + 
  labs(x = "Target of agreement",
       y = "Proportion",
       title = "Proportion of syntactic and semantic agreements in types of targets of agreement") +
  scale_fill_manual(name = "Agreement",
                    labels = c("Syntactic", "Semantic"),
                    values = wes_palette(n=2, name="Darjeeling1")) + 
  scale_x_discrete(labels = c("Attribute", "Predicate", "Relative pronoun")) -> NOAPPsemXtargets
NOAPPsemXtargets

agrs %>% 
  filter(!(order %in% c("AST", "ATS", "TAS"))) %>% 
  ggplot(aes(x = distance, fill = semantic_agr)) + 
  geom_bar(position = "fill") + 
  labs(x = "Agreement",
       y = "Proportion",
       title = "Counts of syntactic and semantic agreements in different distances between target and source") +
  scale_fill_discrete(name = "Agreement",
                      labels = c("Syntactic", "Semantic"))

agrs %>% 
  filter(!(order %in% c("AST", "ATS", "TAS"))) %>% 
  ggplot(aes(x = target_first, fill = semantic_agr)) + 
  geom_bar(position = "fill") + 
  labs(x = "Target-source order",
       y = "Proportion",
       title = "Proportion of syntactic and semantic agreements possible target-source orderings") +
  scale_fill_discrete(name = "Agreement",
                      labels = c("Syntactic", "Semantic")) +
  scale_x_discrete(labels = as_labeller(c(`0` = "Controller first", 
                                          `1` = "Target first")))

agrs %>% 
  filter(!(order %in% c("AST", "ATS", "TAS"))) %>% 
  ggplot(aes(x = source_cat, fill = semantic_agr)) + 
  geom_bar(position = "fill") + 
  labs(x = "Type of noun",
       y = "Proportion",
       title = "Proportion of syntactic and semantic agreements in split and lexical hybrids") +
  scale_fill_discrete(name = "Agreement",
                      labels = c("Syntactic", "Semantic")) +
  scale_x_discrete(labels = as_labeller(c(`depreciative` = "Depreciative", 
                                          `gendered` = "Gendered",
                                          `profession` = "Profession")))
```
Very different results, but I removed a the majority of profession noun occurrences, which I didn't have many anyways. I just need to remember to include order as a factor in my models.

Facet grid just to better see how the order affects proportions for different targets.
```{r}
agrs %>% 
  ggplot(aes(x = order, fill = semantic_agr)) + 
  geom_bar(position = "fill") + 
  facet_grid(rows = vars(type_of_agreement)) + 
  labs(x = "Order",
       y = "Proportion",
       title = "Proportions of syntactic and semantic agreements depending on the order and the target of agreement") +
  scale_fill_discrete(name = "Agreement",
                      labels = c("Syntactic", "Semantic"))
```
Some apparent differences for types of agreement can be explained by ordering with profession nouns.

I will mutate a new factor that tells me if apposition appears before the controller.
```{r}
agrs %>% 
  mutate(appos_first = as.factor(ifelse(order %in% c("AST", "ATS", "TAS"), 1, 0))) -> agrs
```

I need a table of occurrences of semantic agreement with these excluded.
```{r}
agrs %>% 
  filter(appos_first == 0) %>% 
  mutate(semantic_agr = as.numeric(levels(semantic_agr))[semantic_agr],
         type_of_agreement = as.character(type_of_agreement)) %>% 
  filter(source_cat == "profession") %>% 
  group_by(type_of_agreement) %>% 
  summarise(n = n(),
            sem_n = sum(semantic_agr),
            syn_n = n-sem_n) %>% 
  bind_rows(summarise_all(., ~if(is.numeric(.)) sum(.) else "")) %>% 
  mutate(sem_prop = paste0(round(sem_n/n*100, 2), "%"),
         syn_prop = paste0(round(syn_n/n*100, 2), "%")) %>% 
  relocate(sem_prop, .after = sem_n) %>% xtable()
```



I want to cross ordering of target and controller with types of agreement.
```{r}
agrs %>% 
  ggplot(aes(x = type_of_agreement, fill = semantic_agr)) + 
  geom_bar(position = "fill") + 
  facet_grid(rows = vars(target_first), 
             labeller = as_labeller(c(`0` = "Controller first", 
                                      `1` = "Target first"))) + 
  labs(x = "Target of agreement",
       y = "Proportion",
       title = "Proportions of syntactic and semantic agreements depending on the type of agreement and ordering") +
  scale_fill_discrete(name = "Agreement",
                      labels = c("Syntactic", "Semantic")) + 
  scale_x_discrete(labels = c("Attribute", "Predicate", "Relative pronoun"))
```
When the target follows the controller, the likelihood of semantic agreement increases along the Agreement Hierarchy. It's difficult to say what exactly happens when the target is first, because it's impossible to have the target first with relative pronouns, however it doesn't seem to be following Corbett's predictions.

I want to also see proportions of semantic agreement with both types of nouns and ordering.
```{r}
agrs %>% 
  ggplot(aes(x = type_of_agreement, fill = semantic_agr)) + 
  geom_bar(position = "fill") + 
  facet_grid(rows = vars(source_cat),
             cols = vars(target_first), 
             labeller = as_labeller(c(`0` = "Controller first", 
                                      `1` = "Target first",
                                      `depreciative` = "Depreciative forms",
                                      `gendered` = "Gendered nouns",
                                      `profession` = "Profession nouns"))) + 
  labs(x = "Target of agreement",
       y = "Proportion",
       title = "Proportions of syntactic and semantic agreements") +
  scale_fill_discrete(name = "Agreement",
                      labels = c("Syntactic", "Semantic")) + 
  scale_x_discrete(labels = as_labeller(c(`attributive` = "Attributive", 
                                          `predicative` = "Predicate", 
                                          `relpron` = "Relative pronoun")))
```
Any monotonicity I could see in the plot for ordering is overshadowed by the effect of types of nouns.

```{r}
agrs %>% 
  ggplot(aes(x = type_of_agreement, fill = semantic_agr)) + 
  geom_bar(position = "fill") + 
  facet_grid(rows = vars(source_cat),
             labeller = as_labeller(c(`depreciative` = "Depreciative forms",
                                      `gendered` = "Gendered nouns",
                                      `profession` = "Profession nouns"))) + 
  labs(x = "Target of agreement",
       y = "Proportion",
       title = "Proportions of syntactic and semantic agreements in noun types and in types of agreement") +
  scale_fill_manual(name = "Agreement",
                    labels = c("Syntactic", "Semantic"),
                    values = wes_palette(n=2, name="Darjeeling1")) + 
  scale_x_discrete(labels = as_labeller(c(`attributive` = "Attributive", 
                                          `predicative` = "Predicate", 
                                          `relpron` = "Relative pronoun"))) -> semXtargetsXnouns
semXtargetsXnouns
```

```{r}
png(file="../writing/M2/images/semXtargetsXnouns.png", 
    width = 20,
    height = 12,
    units = "cm",
    res = 500)

(semXtargets | semXtargetsXnouns) + plot_layout(guides = "collect", axes = "collect") & labs(title = NULL)

dev.off()
```

```{r}
agrs %>% 
  filter(!(order %in% c("AST", "ATS", "TAS"))) %>% 
  ggplot(aes(x = type_of_agreement, fill = semantic_agr)) + 
  geom_bar(position = "fill") + 
  facet_grid(rows = vars(source_cat),
             labeller = as_labeller(c(`depreciative` = "Depreciative forms",
                                      `gendered` = "Gendered nouns",
                                      `profession` = "Profession nouns"))) + 
  labs(x = "Target of agreement",
       y = "Proportion",
       title = "Proportions of syntactic and semantic agreements in noun types and in types of agreement") +
  scale_fill_manual(name = "Agreement",
                    labels = c("Syntactic", "Semantic"),
                    values = wes_palette(n=2, name="Darjeeling1")) + 
  scale_x_discrete(
    labels = as_labeller(c(`attributive` = "Attributive", 
                           `predicative` = "Predicate", 
                           `relpron` = "Relative pronoun"))) -> NOAPPsemXtargetsXnouns
NOAPPsemXtargetsXnouns
```


```{r}
png(file="../writing/M2/images/NOAPPsemXtargetsXnouns.png", 
    width = 20,
    height = 12,
    units = "cm",
    res = 500)

(NOAPPsemXtargets | NOAPPsemXtargetsXnouns) + plot_layout(guides = "collect", axes = "collect") & labs(title = NULL)

dev.off()
```

What about distance with ordering? Below a plot with proportions of semantic agreement for different distances with target before and after the controller in different types of agreement.
```{r}
(
# left column
agrs %>% 
  mutate(semantic_agr = as.numeric(levels(semantic_agr))[semantic_agr],
         target_first = as.numeric(levels(target_first))[target_first],
         type_of_agreement = as.character(type_of_agreement)) %>% 
  group_by(target_first, type_of_agreement, distance) %>% 
  summarise(prop = sum(semantic_agr)/n()) %>% 
  ungroup() %>% 
  mutate(synt_prop = 1-prop) %>% 
  add_row(target_first = 1, type_of_agreement = "relpron", distance = 0:12, prop = 0) %>% 
  mutate(target_first = as.factor(target_first),
         type_of_agreement = as.factor(type_of_agreement)) %>% 
  filter(target_first == 1) %>% 
  ggplot(aes(x = distance)) +
  geom_col(aes(y = synt_prop+prop, fill = "A")) +
  geom_col(aes(y = prop, fill = "B")) +
  facet_grid(rows = vars(type_of_agreement)) +
  theme(strip.text = element_blank()) +
  scale_x_reverse() + guides(fill = "none")
|
# right column
agrs %>% 
  mutate(semantic_agr = as.numeric(levels(semantic_agr))[semantic_agr],
         target_first = as.numeric(levels(target_first))[target_first],
         type_of_agreement = as.character(type_of_agreement)) %>% 
  group_by(target_first, type_of_agreement, distance) %>% 
  summarise(prop = sum(semantic_agr)/n()) %>% 
  ungroup() %>% 
  mutate(synt_prop = 1-prop) %>% 
  add_row(target_first = 1, type_of_agreement = "relpron", distance = 0:12, prop = 0) %>% 
  mutate(target_first = as.factor(target_first),
         type_of_agreement = as.factor(type_of_agreement)) %>% 
  filter(target_first == 0) %>% 
  ggplot(aes(x = distance)) +
  geom_col(aes(y = synt_prop+prop, fill = "A")) +
  geom_col(aes(y = prop, fill = "B")) +
  facet_grid(rows = vars(type_of_agreement)) + 
  guides(fill = "none")
) + plot_layout(axes = "collect")
```

# Statistical modelling
I need to relevel the factors for statistical modelling.
```{r}
agrs %>% 
  mutate(type_of_agreement = relevel(type_of_agreement, ref = "predicative"),
         target_first = relevel(target_first, ref = "1"),
         source_cat = relevel(source_cat, ref = "depreciative"),
         semantic_agr = relevel(semantic_agr, ref = "0")) -> agrs
```

Up to here was saved as for-modeling.RData

## General model
```{r}
glmer(semantic_agr ~ type_of_agreement + distance + target_first + source_cat + appos_first + (1|type_of_text),
    family = binomial(link = "logit"), 
    data = agrs) -> gen.max.old.m
summary(gen.max.old.m)
```
I realized I should recode appos_first, because it gives 0 to occurrences without any apposition and to occurrences with apposition after the controller, which are actually different.

-1 0 1 instead!
```{r}
agrs %>% 
  mutate(appos_first = as.numeric(ifelse(source_cat == "profession" & appos_first == 0, 0.5, appos_first))) %>% 
  glmer(semantic_agr ~ type_of_agreement + distance + target_first + source_cat + appos_first + (1|type_of_text),
    family = binomial(link = "logit"),
    data = .) -> gen.max.m
summary(gen.max.m)
```
Effect of variables:
1. attributive agreement is less likely to be semantic than predicate; NOT significant
2. relative pronoun agreement is less likely to be semantic than predicate; NOT significant
3. larger distance increases likelihood of semantic agreement; NOT significant
4. target being after controller increases likelihood of semantic agreement; signif. at p<0.001
5. socially gendered nouns are less likely to have semantic agreement than depreciative forms; significant at p<0.001
6. profession nouns are more likely to have semantic agreement than depreciative forms; significant at p<0.001
7. apposition appearing before the controller of agreement increases the likelihood of semantic agreement; significant at p<0.001

4 and 6 aligns with the literature
5 doesn't align with the literature
7 aligns with my personal predictions when I was coding this

For the purpose of finding the smallest fitting model I need to compute a model without random effects.
```{r}
glm(semantic_agr ~ type_of_agreement + distance + target_first + source_cat + appos_first,
    family = binomial(link = "logit"), 
    data = agrs %>% 
  mutate(appos_first = as.numeric(ifelse(source_cat == "profession" & appos_first == 0, 0.5, appos_first)))) -> gen.nore.m
summary(gen.nore.m)
```
It's very similar to the model with random effects!

Now I want to see what's the best model for this data.
```{r}
step(gen.nore.m, direction = "backward") -> gen.bw.m
gen.bw.m
```

try loo_compare for model comparisons

glmercontrol optimex package for singulatrity
So the best model for this data should be the following:
```{r}
agrs %>% 
  mutate(appos_first = as.numeric(ifelse(source_cat == "profession" & appos_first == 0, 0.5, appos_first))) %>% 
glmer(semantic_agr ~ target_first + source_cat + appos_first + (1|type_of_text),
    family = binomial(link = "logit"), 
    data = .) -> gen.min.m
summary(gen.min.m)
```
Let's just compare the maximum and minimum models.
```{r}
anova(gen.max.m, gen.min.m)
```
I suppose it's not too much better. For now I'm gonna only report the maximum model.

Trying to plot the results of the model, but I don't understand this fully for now. (OR is odds ratio, so maybe I can figure this out!)
```{r}
ggstats::ggcoef_model(gen.max.m, exponentiate = T)
```

## Profession nouns model
```{r}
agrs %>% 
  filter(source_cat == "profession") %>% 
  glm(semantic_agr ~ type_of_agreement + distance + target_first + appos_first, 
      family = binomial(link = "logit"), data = .) -> prof.m
summary(prof.m)
```
Effect of variables:
1. attributive agreement more likely to be semantic than predicate agreement; signif. at p<0.05
2. relative pronoun agreement more likely to be semantic than predicate agreement; significant at p<0.05
3. larger distance increases likelihood of semantic agreement; NOT significant
4. target after controller increases likelihood of semantic agreement; NOT significant
5. apposition before the controller increases likelihood of semantic agreement; signif. at p<0.001

2 aligns with the literature
1 doesn't align with the literature
5 aligns with my personal predictions when I was coding this

I should see if this model is better or the one based on order instead of appos_first.
```{r}
agrs %>% 
  filter(source_cat == "profession") %>% 
  glm(semantic_agr ~ type_of_agreement + distance + target_first + order, 
      family = binomial(link = "logit"), data = .) -> prof.ord.m
summary(prof.ord.m)
```
This one has one coefficient not defined because of singularities, so I assume this one is worse.

## Gendered nouns model
```{r}
agrs %>% 
  filter(source_cat == "gendered") %>% 
  glm(semantic_agr ~ type_of_agreement + distance + target_first, 
      family = binomial(link = "logit"), data = .) -> gend.m
summary(gend.m)
```
Effect of variables:
1. attributive agreement more likely to be semantic than predicate agreement; NOT significant
2. relative pronoun agreement less likely to be semantic than predicate agreement; NOT significant
3. larger distance increases likelihood of semantic agreement; significant at p<0.05
4. target after controller increases likelihood of semantic agreement; NOT significant

Kind of useless, the only significant predictor is the distance.

## Depreciative forms model
```{r}
agrs %>% 
  filter(source_cat == "depreciative") %>% 
  glm(semantic_agr ~ type_of_agreement + distance + target_first, 
      family = binomial(link = "logit"), data = .) -> depr.m
summary(depr.m)
```
Effect of variables:
1. attributive agreement less likely to be semantic than predicate agreement; NOT significant
2. relative pronoun agreement less likely to be semantic than predicate agreement; NOT significant
3. larger distance decreases likelihood of semantic agreement; NOT significant
4. target after controller increases likelihood of semantic agreement; significant at p<0.001

The only significant predictor is the target-controller ordering: target after controller increases likelihood of semantic agreement (significant at p<0.001).

# Experimental results
Loading data
```{r}
list.files("./results/depr/", pattern = "*.csv") -> filenames
ldf <- lapply(paste0("./results/depr/", filenames), read.csv)
rbindlist(ldf) -> depr
depr %>% 
  mutate(study = "depr") -> depr

list.files("./results/prof/", pattern = "*.csv") -> filenames
ldf <- lapply(paste0("./results/prof/", filenames), read.csv)
rbindlist(ldf) -> prof
prof %>% 
  mutate(study = "prof") -> prof

exper <- rbind(depr, prof)
rm(filenames, ldf, depr, prof)
```

## Demographic data
Getting the demographic data in a separate dataframe.
```{r}
exper %>% 
  filter(grepl("age", response)) %>% 
  select(id, subject_id, study, response) %>% 
  rowwise() %>% 
  mutate(age = as.numeric(str_extract(response, '(?<=age\":\")\\d+')),
         gender = str_extract(response, '(?<=gender\":\").+(?=\",\"languages)'),
         languages = str_extract(response, '(?<=languages\":\").+(?=\",\")'),
         country = str_extract(response, '(?<=country\":\").+(?=\")')) %>% 
  mutate(gender = case_when(str_starts(gender, "m|M") ~ "m",
                            str_starts(gender, "k|K|ż") ~ "f"),
         languages = ifelse(str_detect(languages, "^\\w+$", negate = T), 
                            languages,
                            ifelse(str_detect(languages, "pol|Pol"), "PL", languages)),
         country = ifelse(str_detect(country, "^\\w+$"), 
                          ifelse(str_detect(country, "pol|Pol"), "PL", country),
                          ifelse(str_detect(country, "W Pol"), "PL", country))) -> demo
```

How many participants didn't speak only Polish or grew up in places other than Poland?
```{r}
demo %>% 
  filter(languages != "PL" | country != "PL")
```
Very few of them, I can filter them out.
```{r}
demo %>% 
  filter(languages != "PL" | country != "PL") %>% 
  select(id) -> to_rem

exper %>% 
  filter(!(id %in% to_rem$id)) -> exper

demo %>% 
  filter(!(id %in% to_rem$id)) -> demo
rm(to_rem)
```

The demographic data looks fine now, I can remove some columns.
```{r}
demo %>% 
  select(id, study, age, gender) -> demo
```

Recoding ages to have groups instead of exact numbers.
```{r}
demo %>% 
  mutate(age = ifelse(age < 31, "18-30", ifelse(age > 50, "51+", "31-50"))) -> demo
```

## Cleaning the main dataframe
Creating a separate dataframe for comprehension question responses, so that I can include them in the full dataframe later.
```{r}
exper %>% 
  filter(condition != "" & trial_type == "html-button-response") %>% 
  select(id, trial_index, response, correct) %>% 
  mutate(CQid = id,
         CQtrial_index = trial_index,
         response = ifelse(response == 0, 1, ifelse(response == 1, 0, response)),
         correct = case_when(correct == "tak" ~ 1,
                             correct == "nie" ~ 0),
         CQcorrect = ifelse(response == correct, 1, 0)) %>% 
  select(!c(response, correct, id, trial_index)) -> CQresponses
```

Removing the rows without acceptability judgements, removing unnecessary columns, mutating new columns and rearranging them.
```{r}
exper %>% 
  filter(condition != "" & trial_type == "survey-likert") %>% 
  mutate(rating = as.numeric(str_extract(response, "\\d(?=\\})"))+1,
         trial_type = as.factor(ifelse(condition == "null", 
                                       "distractor", 
                                       ifelse(item == "null",
                                              "control",
                                              "experimental"))),
         agreement = as.factor(case_when(str_starts(condition, "f") & study == "prof" ~ "semantic",
                                         str_starts(condition, "m") & study == "prof" ~ "syntactic",
                                         str_starts(condition, "f") & study == "depr" ~ "syntactic",
                                         str_starts(condition, "m") & study == "depr" ~ "semantic")),
         condition = as.factor(ifelse(trial_type == "experimental", 
                                      str_extract(condition, "(?<=-)\\w+"), 
                                      condition)),
         CQcorrect = (CQresponses %>% filter(CQid == id & CQtrial_index == trial_index+1))$CQcorrect,
         noun = as.factor(study),
         id = as.factor(id),
         group = as.factor(group),
         item = as.factor(item),
         subject_id = as.factor(subject_id)) %>% 
  select(time_elapsed, id, trial_index, trial_type, group, item, condition, 
         agreement, noun, rating, rt, CQcorrect, stimulus, subject_id) -> exper
rm(CQresponses)
```

Add demo back to exper
```{r}
exper %>% 
  rowwise() %>% 
  mutate(age = demo[which(demo$id == id), ]$age,
         gender = demo[which(demo$id == id), ]$gender) -> exper
```

## Descriptive stats
Percentage of correct answers to CQs per participant
```{r}
exper %>% 
  group_by(id) %>% 
  summarise(QCscore = paste0(round(sum(CQcorrect, na.rm = T)/60*100, 2), "%"))
```

Average acceptability ratings of control items.
```{r}
exper %>% 
  filter(trial_type == "control") %>% 
  group_by(condition) %>% 
  summarise(mean(rating))
```

Did all participants rate ungrammatical below grammatical?
```{r}
exper %>% 
  filter(trial_type == "control") %>% 
  group_by(id, condition) %>% 
  summarise(m_rating = mean(rating)) %>% 
  tidyr::spread(condition, m_rating) %>% 
  mutate(gram_rating = correct,
         ungram_rating = incorrect,
         correct = ifelse(gram_rating > ungram_rating, 1, 0)) %>% 
  select(!(incorrect)) %>% 
  filter(correct == 0)
```
Yes!

Did any of the participants rate all of the sentences with depreciative forms 1, but other items higher?
```{r}
exper %>% 
  filter(noun == "depr") %>% 
  group_by(id, trial_type) %>% 
  summarise(min = min(rating), avg = mean(rating), max = max(rating)) %>% 
  filter(trial_type == "experimental" & max == 1)

exper %>% 
  filter(noun == "depr") %>% 
  filter(id == "6dly03" | id == "wvb2ar") %>% 
  group_by(id, trial_type) %>% 
  summarise(min = min(rating), avg = mean(rating), max = max(rating))
```
They rated all my depreciative forms 1!!! what do i do!? exclude!
```{r}
exper %>% filter(!(id == "6dly03" | id == "wvb2ar")) -> exper
demo %>% filter(!(id == "6dly03" | id == "wvb2ar")) -> demo
```

Up to here is saved as exp-cleaned.RData

Final demographic profile of participants.
```{r}
demo %>% 
  filter(study == "prof") %>% 
  group_by(gender, age) %>% 
  count() %>% tidyr::spread(age, n) %>% xtable()

demo %>% 
  filter(study == "depr") %>% 
  group_by(gender, age) %>% 
  count() %>% tidyr::spread(age, n) %>% xtable()
```

Average acceptability ratings per experiment and per condition.
```{r}
exper %>%
  filter(trial_type == "experimental") %>% 
  group_by(noun, condition, agreement) %>% 
  summarise("mean rating" = mean(rating),
            sd = sd(rating),
            first = quantile(rating, .25),
            median = median(rating),
            third = quantile(rating, .75),
            mode = mlv(rating, method="mfv"))

exper %>%
  filter(trial_type == "experimental") %>% 
  mutate(condition = factor(condition),
         agreement = factor(agreement, 
                            levels = c("syntactic", "semantic"),
                            labels = c("Syntactic", "Semantic"))) %>% 
  group_by(condition, noun, agreement) %>% 
  summarise(sd = sd(rating),
            rating = mean(rating)) %>% 
  ggplot(aes(x = condition, y = rating, fill = agreement)) +
  geom_col(position = "dodge") +
  geom_errorbar(aes(ymin = rating - sd,
                    ymax = rating + sd),
                width=.5,
                position=position_dodge(.9)) +
  facet_grid(cols = vars(factor(noun, levels = c("prof", "depr"))),
             labeller = as_labeller(c(`prof` = "Exp. 1 (profession nouns)",
                                      `depr` = "Exp. 2 (depreciative forms)"))) +
  scale_x_discrete(name = "Condition",
                   limits = c("pre", "post", "pron"),
                   labels = c("Prenominal\npredicate", "Postnominal\npredicate", "Relative\npronoun")) +
  scale_y_discrete(name = "Rating",
                   limits = c("1", "2", "3", "4", "5")) +
  scale_fill_manual(name = "Agreement", labels = c("Syntactic", "Semantic"),
                    values = wes_palette(n=2, name="Darjeeling1")) -> main.plot
main.plot
```
Tables above for latex:
```{r}
exper %>%
  filter(trial_type == "experimental" & noun == "prof") %>% 
  group_by(condition, agreement) %>% 
  summarise("mean rating" = mean(rating),
            sd = sd(rating)) %>% 
  tidyr::pivot_wider(names_from = condition,
                     values_from = c(`mean rating`, sd),
                     names_vary = "slowest") %>% xtable()

exper %>%
  filter(trial_type == "experimental" & noun == "depr") %>% 
  group_by(condition, agreement) %>% 
  summarise("mean rating" = mean(rating),
            sd = sd(rating)) %>% 
  tidyr::pivot_wider(names_from = condition,
                     values_from = c(`mean rating`, sd),
                     names_vary = "slowest") %>% xtable()
```


A plot with the ratings of control items, just to compare with the ratings of experimental items:
```{r}
exper %>%
  filter(trial_type == "control") %>% 
  mutate(condition == factor(condition)) %>% 
  group_by(condition, agreement, trial_type) %>% 
  summarise(sd = sd(rating),
            rating = mean(rating)) %>% 
  ggplot(aes(x = condition, y = rating, fill = trial_type)) +
  geom_col(position = "dodge") +
  geom_errorbar(aes(ymin = rating - sd,
                    ymax = rating + sd),
                width=.5,
                position=position_dodge(.9)) +
  facet_grid(cols = vars(trial_type),
             labeller = as_labeller(c(`control` = "Control"))) +
  scale_x_discrete(name = "Condition",
                   limits = c("incorrect", "correct"),
                   labels = c("Ungram.", "Gram.")) +
  scale_y_discrete(name = "Rating",
                   limits = c("1", "2", "3", "4", "5")) +
  scale_fill_manual(values = wes_palette(n=3, name="Darjeeling1")[3]) +
  guides(fill = "none") -> contr.plot
contr.plot
```

Putting the plots together:
```{r}
png(file="../writing/M2/images/expmainplot.png", 
    width = 22,
    height = 8,
    units = "cm",
    res = 500)

guide_area() + main.plot + contr.plot + plot_layout(guides = "collect",
                                                    axes = "collect",
                                                    widths = c(1,10,2))

dev.off()
```

Maybe I want a small plot showing the differences between syntactic and semantic agreement, I will first make a table here.
```{r}
exper %>%
  filter(trial_type == "experimental") %>% 
  group_by(noun, condition, agreement) %>% 
  summarise(avg_rating = mean(rating)) %>% 
  tidyr::pivot_wider(names_from = agreement, values_from = avg_rating) %>% 
  summarise(diff = semantic - syntactic)

exper %>%
  filter(trial_type == "experimental") %>% 
  group_by(noun, condition, agreement) %>% 
  summarise(avg_rating = mean(rating)) %>% 
  tidyr::pivot_wider(names_from = agreement, values_from = avg_rating) %>% 
  summarise(diff = semantic - syntactic) %>% 
  mutate(noun = factor(noun, levels = c("prof", "depr"))) %>% 
  ggplot(aes(x = condition, y = diff, fill = noun)) +
  geom_col(position = "dodge") +
  scale_x_discrete(name = "Condition",
                   limits = c("pre", "post", "pron"),
                   labels = c("Prenominal\npredicate", "Postnominal\npredicate", "Relative\npronoun")) +
  scale_y_continuous(name = "Difference between semantic/syntactic rating") +
  scale_fill_manual(name = "Noun types",
                    limits = c("prof", "depr"),
                    labels = c("Profession nouns", "Depreciative forms"),
                    values = wes_palette(n=5, name="Darjeeling1")[4:5]) -> diffs.plot
diffs.plot
```
Saving this plot:
```{r}
png(file="../writing/M2/images/expdiffsplot.png", 
    width = 16,
    height = 12,
    units = "cm",
    res = 500)

diffs.plot

dev.off()
```

### Plots with demographic information
Plot showing the between gender differences in judgements.
```{r}
exper %>%
  filter(noun == "prof" & trial_type == "experimental") %>% 
  group_by(condition, agreement, gender) %>% 
  summarise(rating = mean(rating)) %>% 
  ggplot(aes(x = gender, y = rating, fill = agreement)) +
  geom_col(position = "dodge") +
  facet_grid(cols = vars(factor(condition, levels = c("pre", "post", "pron"))), 
             labeller = as_labeller(c(`pre` = "Prenominal predicate",
                                      `post` = "Postnominal predicate",
                                      `pron` = "Relative pronoun"))) +
  scale_y_discrete(limits = c("1", "2", "3", "4", "5"))

exper %>%
  filter(noun == "depr" & trial_type == "experimental") %>% 
  group_by(condition, agreement, gender) %>% 
  summarise(rating = mean(rating)) %>% 
  ggplot(aes(x = gender, y = rating, fill = agreement)) +
  geom_col(position = "dodge") +
  facet_grid(cols = vars(factor(condition, levels = c("pre", "post", "pron"))), 
             labeller = as_labeller(c(`pre` = "Prenominal predicate",
                                      `post` = "Postnominal predicate",
                                      `pron` = "Relative pronoun"))) +
  scale_y_discrete(limits = c("1", "2", "3", "4", "5"))
```
PROF NOUNS:
1. For prenominal predicates (Zgromadzonych przywitał/a burmistrz) men preferred syntactic agreement (przywitał), while women slightly preferred semantic agreement (przywitała)
2. For postnominal predicates (Burmistrz przywitał/a zgromadzonych) both genders prefered semantic agreement (przywitała), but for women there was a bigger difference.

DEPR FORMS:
1. Both genders preferred semantic agreement in all conditions, with the exception of women with relative pronoun agreement, where they gave higher ratings to syntactic agreement (Reżysery, które...)
```{r}
exper %>%
  filter(noun == "depr" & trial_type == "experimental") %>% 
  group_by(condition, agreement, gender) %>% 
  summarise(rating = mean(rating)) %>% 
  ggplot(aes(x = condition, y = rating, fill = agreement)) +
  geom_col(position = "dodge") +
  facet_grid(cols = vars(gender)) +
  scale_x_discrete(limits = c("pre", "post", "pron")) +
  scale_y_discrete(limits = c("1", "2", "3", "4", "5"))
```
2. For men especially big difference between acceptability of postnominal predicates: semantic agreement a lot better.

Plot with age groups
```{r}
demo %>% 
  filter(study == "prof") %>% 
  group_by(age) %>% 
  count()

exper %>%
  filter(noun == "prof" & trial_type == "experimental") %>% 
  group_by(condition, agreement, age) %>% 
  summarise(rating = mean(rating)) %>% 
  ggplot(aes(x = condition, y = rating, fill = agreement)) +
  geom_col(position = "dodge") +
  facet_grid(cols = vars(age)) +
  scale_x_discrete(limits = c("pre", "post", "pron")) +
  scale_y_discrete(limits = c("1", "2", "3", "4", "5"))

demo %>% 
  filter(study == "depr") %>% 
  group_by(age) %>% 
  count()

exper %>%
  filter(noun == "depr" & trial_type == "experimental") %>% 
  group_by(condition, agreement, age) %>% 
  summarise(rating = mean(rating)) %>% 
  ggplot(aes(x = condition, y = rating, fill = agreement)) +
  geom_col(position = "dodge") +
  facet_grid(cols = vars(age)) +
  scale_x_discrete(limits = c("pre", "post", "pron")) +
  scale_y_discrete(limits = c("1", "2", "3", "4", "5"))
```
PROF NOUNS:
1. For the youngest group syntactic agreement was only preferred with prenominal predicates (Zgromadzonych przywitał burmistrz).
2. For the middle group semantic agreement was preferred in every condition
3. The oldest group is just one person, but it seems she prefers the semantic agreement even more in every condition

DEPR NOUNS:
1. Youngest group prefers semantic agreement across conditions
2. The middle group prefers syntactic agreement only for relative pronouns
3. The oldest group really dislikes the depreciative forms, they only somewhat accept them for prenominal predicates, where they prefer syntactic agreement (Do Ameryki poleciały reżysery)

A plot with both age and gender
```{r}
demo %>% 
  filter(study == "prof") %>% 
  group_by(age, gender) %>% 
  count()

exper %>%
  filter(noun == "prof" & trial_type == "experimental") %>% 
  group_by(condition, agreement, age, gender) %>% 
  summarise(rating = mean(rating)) %>% 
  ggplot(aes(x = condition, 
             y = rating, 
             fill = factor(agreement, 
                           levels = c("syntactic", "semantic"),
                           labels = c("Syntactic", "Semantic")))) +
  geom_col(position = "dodge") +
  facet_grid(cols = vars(age),
             rows = vars(gender),
             labeller = as_labeller(c(`18-30` = "18-30",
                                      `31-50` = "31-50",
                                      `51+` = "51+",
                                      `f` = "Women",
                                      `m` = "Men"))) +
  scale_x_discrete(name = "Condition",
                   limits = c("pre", "post", "pron"),
                   labels = c("Prenominal\npredicate", 
                              "Postnominal\npredicate", 
                              "Relative\npronoun")) +
  scale_y_discrete(name = "Rating",
                   limits = c("1", "2", "3", "4", "5")) +
  scale_fill_manual(name = "Agreement",
                    values = wes_palette(n=2, name="Darjeeling1")) -> demo.prof.plot
demo.prof.plot

demo %>% 
  filter(study == "depr") %>% 
  group_by(age, gender) %>% 
  count()

exper %>%
  filter(noun == "depr" & trial_type == "experimental") %>% 
  group_by(condition, agreement, age, gender) %>% 
  summarise(rating = mean(rating)) %>% 
  ggplot(aes(x = condition, 
             y = rating, 
             fill = factor(agreement, 
                           levels = c("syntactic", "semantic"),
                           labels = c("Syntactic", "Semantic")))) +
  geom_col(position = "dodge") +
  facet_grid(cols = vars(age),
             rows = vars(gender),
             labeller = as_labeller(c(`18-30` = "18-30",
                                      `31-50` = "31-50",
                                      `51+` = "51+",
                                      `f` = "Women",
                                      `m` = "Men"))) +
  scale_x_discrete(name = "Condition",
                   limits = c("pre", "post", "pron"),
                   labels = c("Prenominal\npredicate", 
                              "Postnominal\npredicate", 
                              "Relative\npronoun")) +
  scale_y_discrete(name = "Rating",
                   limits = c("1", "2", "3", "4", "5")) +
  scale_fill_manual(name = "Agreement", 
                    values = wes_palette(n=2, name="Darjeeling1")) -> demo.depr.plot
demo.depr.plot
```
PROF NOUNS:
1. The youngest women prefer syntactic agreement in every condition, but older always prefer semantic (could be that younger women would prefer to use feminitives, while older women dont care for them that much)
2. For men syntactic prenominal predicate agreement is either the same or better than semantic, while for other conditions they prefer semantic agreement. This is in line with Corbett's prediction that target before controller is more likely to be syntactic.

DEPR FORMS:
1. For relative pronouns women in both groups prefer syntactic agreement, but for predicates always semantic.
2. Men mostly prefer semantic agreement across age groups and conditions, with the exception of relative pronouns for men aged 31-50, where syntactic agreement is slightly more preferred. 
3. The oldest is just one man.

Saving the 2 plots above.
```{r}
png(file="../writing/M2/images/expdemoprof.png", 
    width = 22,
    height = 9,
    units = "cm",
    res = 500)

demo.prof.plot

dev.off()

png(file="../writing/M2/images/expdemodepr.png", 
    width = 22,
    height = 9,
    units = "cm",
    res = 500)

demo.depr.plot

dev.off()
```


Boxplots, but I'm not sure if they are very interesting.
```{r}
exper %>%
  filter(trial_type == "experimental") %>% 
  group_by(condition, noun, agreement) %>% 
  ggplot(aes(x = condition, y = rating, fill = agreement)) +
  geom_boxplot(outlier.shape = NA) +
  # geom_jitter(aes(color = as.factor(rating)), width = .2) +
  geom_point(position = position_jitter(height=0,width = .1),alpha=.4,size=1.5) + # from Emma
  scale_color_manual(values = c("1" = rgb(0, 255-51*0, 0, maxColorValue = 255),
                                "2" = rgb(0, 255-51*1, 0, maxColorValue = 255),
                                "3" = rgb(0, 255-51*2, 0, maxColorValue = 255),
                                "4" = rgb(0, 255-51*3, 0, maxColorValue = 255),
                                "5" = rgb(0, 255-51*4, 0, maxColorValue = 255))) +
  facet_grid(cols = vars(noun))
```

## Statistical modeling
Data for modeling: 
```{r}
exper %>% 
  filter(trial_type == "experimental") %>% 
  ungroup() %>% # I have no idea why this is necessary, but otherwise I can't have as.ordered(rating)
  mutate(order = factor(case_when(condition == "pre" ~ "TargetFirst", 
                                  condition == "post" ~ "ControllerFirst", 
                                  condition == "pron" ~ "ControllerFirst"), 
                        levels = c("TargetFirst", "ControllerFirst")),
         target_type = factor(ifelse(condition == "pron", "RelPron", "Pred"), 
                              levels = c("Pred", "RelPron")),
         group = as.factor(group)) %>%
  select(id, trial_index, time_elapsed, group, item, order, target_type, agreement, noun, rating, age, gender) %>% 
  mutate(order = relevel(order, ref = "TargetFirst"),
         target_type = relevel(target_type, ref = "Pred"),
         agreement = relevel(agreement, ref = "syntactic"),
         noun = relevel(noun, ref = "depr"),
         age = as.factor(age),
         gender = as.factor(gender),
         rating = as.ordered(rating)) -> mod
```

Up to here saved as exp-modeling.RData

A general model with data from both experiments:
```{r}
mod %>% 
  clmm(rating ~ noun*agreement*(order + target_type + (1|id) + (1|item) + (1|trial_index) + (1|age) + (1|group)),
       data = .) -> model
summary(model)
```
1. Profession noun sentences had significantly higher ratings than depreciative forms (duh, p<0.001)

### Profession nouns
Model!
```{r}
mod %>% 
  filter(noun == "prof") %>% 
  clmm(rating ~ agreement*(order + target_type + (1|id) + (1|item) + (1|trial_index) + (1|age) + (1|group)),
       data = .) -> prof.mod
summary(prof.mod)
```
1. Relative pronouns are significantly more acceptable than predicates (p<0.05)

Let's try some model comparisons!
```{r}
clm(rating ~ agreement*(order + target_type),
    data = mod %>% filter(noun == "prof")) -> prof.mod.nore
step(prof.mod.nore, direction = "backward") -> prof.bw
```
Which is best?
```{r}
prof.bw
```
I'm running a model like this then, only with random effects added.
```{r}
mod %>% 
  filter(noun == "prof") %>% 
  clmm(rating ~ agreement + order + target_type + agreement:order + (1|id) + (1|item) + (1|trial_index) + (1|age) + (1|group),
       data = .) -> prof.mod.best
summary(prof.mod.best)
```
I'm not sure if it makes a lot of sense to report this model, I will stick with the original one.

### Depreciative forms
Model!
```{r}
mod %>% 
  filter(noun == "depr") %>% 
  clmm(rating ~ agreement*(order + target_type + (1|id) + (1|item) + (1|trial_index) + (1|age) + (1|group)),
       data = .) -> depr.mod
summary(depr.mod)
```
No significant terms <3

Let's try some model comparisons!
```{r}
clm(rating ~ agreement*(order + target_type),
    data = mod %>% filter(noun == "depr")) -> depr.mod.nore
step(depr.mod.nore, direction = "backward") -> depr.bw
```
Which is best?
```{r}
depr.bw
```
I'm running a model like this then, only with random effects added.
```{r}
mod %>% 
  filter(noun == "depr") %>% 
  clmm(rating ~ agreement + (1|id) + (1|item) + (1|trial_index) + (1|age) + (1|group),
       data = .) -> depr.mod.best
summary(depr.mod.best)
```
This seems completely useless.





















